---
title: "HW.Transform.1"
author: "Logan Hardy"
date: "01/19/2022"
output: html_document
---

30 points (25 for this assignment, 5 for being on time)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

# Step 1: Inequalities (5 points)
```{r}
num1 <- 150
num2 <- 50
num3 <- 100
num4 <- "150"

# Write an inequality to see if num1 or num2 are greater than num3
num1 | num2 > num3

# Write an inequality to see if both num1 and num2 are greater than num3
num1 & num2 > num3

# Write an inequality to check if num2 is both less than or equal to num1 and greater than or equal to num3
(num2 <= num1) & (num2 >= num3)

# Write an inequality to check if num2 + num3 is NOT equal to num1
num2 + num3 != num1

# Write an inequality to see if num1 and num4 are equal
num1 == num4

```

# Step 2: Mathematical Operations (2 points)
```{r}
# Multiply num1 and num2 and divide that result by num3. Just return the remainder.
(num1 * num2) %/% num3


```

Q: In your console, try using num4 in a mathematical operation. Does this work? Why or why not? 
A: No, this does not work because it is a string.  Error in num4 + 20 : non-numeric argument to binary operator

# Step 3: (4 points)
```{r}
vec1 <- c(1:4)
vec2 <- c(45, 15, 2, 11)
vec3 <- c(64, 2, 199, 358, 34, 7)

# Divide vec2 by vec1 and get the answer in an integer. 
as.integer(vec2 / vec1)

# Write an inequality to see if vec2 is greater than or equal to vec1 ans save the results as vec4.
vec4 <- vec2 >= vec1

#vec3 / vec2

```


Q: What would happen if you divided vec3 by vec2? 
A: It would work by repeating through the list but give a warning because of the different list lengths. Warning in vec3/vec2 : longer object length is not a multiple of shorter object length

Q: Try a mathematical operation with vec4 in the console. What happens? 
A: Mathematical operation works with vec4 because logic variables are treated as 1's and 0's.  So if you have vec4 + 10 the result would be: 11 11 10 11.  

# Step 4: Import Data (2 points)
```{r}
# Import the hazmat77_14 dataset with readr. 
hazmat77_14 <- read_csv("../Data/hazmat77_14.csv")

```

I recommend you run some exploration functions quickly in the console to get to know this data. 

Q: If you wanted to quickly see the mean, median, and quartiles of all of the numeric variables in this dataset, which exploration function would you use? 
A: summary(); for example: summary(hazmat77_14)

# Step 5: Filter, Select, Mutate, etc. (12 points)

Follow these instructions carefully, step by step. (I suggest you try to do these one at a time so you can catch errors as you write them.)

Create a subset of the hazmat77_14 dataset by using the following steps in order: 

1. Filter for the following specifications: 
    a) the date is after the 1st of January, 2000 
    b) the only report submission source is web and 
    c) the origin state is Utah
2. Select the date column, and all columns that are numeric.
3. Create a new column called Total.Fatalities.Injuries that is Total.Hazmat.Fatalities, Total.Hazmat.Hosp.Injuries, and Total.Hazmat.NonHosp.Injuries added together. 
4. Filter this new dataset so that the new column you created is greater than 0. 

```{r}
hazmat77_14 %>% filter(Date.of.Incident > "2000-01-01", 
                       Report.Submission.Source == "Web",
                       Origin.State == "UTAH") %>% select(Date.of.Incident, is.numeric) %>% mutate(Total.Fatalities.Injuries = Total.Hazmat.Fatalities + Total.Hazmat.Hosp.Injuries + Total.Hazmat.NonHosp.Injuries) %>% filter(Total.Fatalities.Injuries > 0)
```

Q: What is the date of the incident with 11 total injuries or fatalities? 
A: 2005-08-10	

Q: How many rows are remaining?
A: 4

Create another subset of the hazmat77_14 dataset by using the following steps in order: 

1. Only include rows where: 
    a) The incident happened in the cities of Ogden, West Valley, or Vernal
    b) The hazardous class was the flammable, combustible liquid category
    c) The damages amount was greater than 0 and less than 1000
2. Create a new column called Damages that would say for the second row: "The flammable - combustiable liquid that came from Utah to Vernal on 2002-10-16 via highway cost a total of $75."
*Note*: you can ignore capitals if you want, but brownie points if you can make the sentence read exactly as above. 
3. Remove all rows with NAs
4. Select just the columns you needed to create the new column and the new column itself. 
5. Limit the results to ten rows with head()


```{r}
hazmat77_14 %>% filter(Incident.City %in% c("OGDEN", "WEST VALLEY", "VERNAL"), 
                       Hazardous.Class == "FLAMMABLE - COMBUSTIBLE LIQUID", 
                       Total.Amount.of.Damages > 0 & Total.Amount.of.Damages < 1000) %>% 
  mutate(Damages = paste0("The ", str_to_lower(Hazardous.Class), " that came from ", str_to_sentence(Origin.State), " to ", str_to_title(Incident.City), " on ", Date.of.Incident, " via ", Mode.of.Transportation, " cost a total of $", Total.Amount.of.Damages, ".")) %>% 
  drop_na() %>% 
  select(Origin.State, Incident.City, Date.of.Incident, Mode.of.Transportation, Hazardous.Class, Total.Amount.of.Damages, Damages) %>% head(10)
```             

Q: How many of these remaining observations came from California? 
A: 3

# Optional feedback: 

How hard was this assignment? Was it: too hard, too easy, or just right? 

ANSWER: Just Right

Approximately how long did this assignment take you? 

ANSWER: 2.5 hours
