---
title: "HW.Explore.2"
author: "Logan Hardy"
date: "1/11/2022"
output: html_document
---

20 points (15 for this assignment, 5 for being on time)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Import the data (2 points)
Import the data using the third method we used in the in class file (readr with character to factor conversion) 

*Optional* To get the date time column (Timestamp) in the right format, add a comma after the filepath and paste in the code: 
col_types = cols(Timestamp = col_datetime(format = "%Y-%m-%d %H:%M:%S")) 
before the closing parenthesis. 

```{r}
mental_health_tech_survey <- read_csv("../Data/mental_health_tech_survey.csv", col_types = cols(Timestamp = col_datetime(format = "%Y-%m-%d %H:%M:%S"))) %>% mutate_if(is.character, as.factor)


```

# Explore the data using the view tab (1 point)
Click on the tab in RStudio that says mental_health_tech_survey.

Q: Based on what you see, which column(s), if any, are unstructured data? 
A: Comments

# Explore the data with str() (3 points)
```{r}
str(mental_health_tech_survey)
```

Q: How many rows and columns are there? 
A: Rows: 1259, columns: 27

Q: What level of measurement is the no_employees column? (This stands for number of employees.)
A: Factor with 6 levels 

Q: What would be the appropriate data type for this column?
A: Order factor 

# Explore the data with head() (2 points)
Use head to show you the top ten rows. (you will have to use n=)
```{r}
head(mental_health_tech_survey, n=10)
```

Q: How many countries are represented in the first ten rows of data? 
A: 3

Q: One of the respondents in this first ten rows works at a company with more than 1000 employees. What is the gender of this person? 
A: M (Male)

# Explore the data with summary() (5 points)
```{r}
summary(mental_health_tech_survey)
```

Q: How many NAs are in the comments column? 
A: 1096

*Note*: The column "supervisor" is a response to the question: Would you be willing to discuss a mental health issue with your direct supervisor(s)? The "coworkers" column similarly answers "Would you be willing to discuss a mental health issue with your coworkers?".  

Q: Are fewer people willing to discuss a mental health issue with their supervisor or coworkers? Compare how many people said "no" to these questions. 
A: supervisors

Q: What is the most common state represented in the survey? 
A: CA (California)

Q: What is the median age in the dataset? (hint: this may show up in scientific notation, which means you may need to move the decimal place to the right.)
A: 31

Q: How many rows of missing data are there in the self employeed column? 
A: 18

# Coerce the data type (2 points)
Coerce the comment column to a character data type. 

```{r}
mental_health_tech_survey$comments <- as.character(mental_health_tech_survey$comments)
```


# Optional feedback: 

How hard was this assignment? Was it: too hard, too easy, or just right? 

ANSWER: Just right

Approximately how long did this assignment take you? 

ANSWER: 2.5 hours
