---
title: "Report.1 In Class Finisher (RMD Tricks)"
author: "Laurel Galli-Graves"
date: "`r format(Sys.time(), '%B %m, %Y')`"
output:
  word_document: 
    toc: yes
  html_document: 
    toc: yes
    toc_float: yes
---

# YAML

1.  Change the date to be the knitting date always with "`r format(Sys.time(), '%B %m, %Y')`"

2.  Table of Contents

-   First look at the table of contents shown in RStudio
-   Look at the code guide
-   Knit the doc and look at the format
-   Now make a table of contents by replacing output: html_Document with the following: output: html_document: toc: yes Knit the doc and look at the results.
-   Make it a floating toc by adding toc_float: yes below it.

# Adjusting the set up chunk

**Remove** messages, warnings, and code.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE, 
                      message = FALSE
                      )
library(tidyverse)
library(ggthemes)
library(pander)
theme_new <-  ggthemes::theme_fivethirtyeight(base_size=12#, base_family="Open Sans"
                                              ) %+replace% theme(panel.grid.major.y = element_line(colour = "grey80", size = 0.25),  panel.grid.major.x = element_blank(), panel.background = element_rect(fill = "white"), plot.background = element_rect(fill = "white"), legend.background = element_rect(fill = "white"))

```

Q: How do you know what packages to include?  
A: Always include tidyverse, then import whatever package is at the top of the finisher file. If you hit a "function not found" google the name of the function and usually there will be the name of the package in the top left corner. 

Q: What does include = FALSE do?  
A: Runs the code but hides both output and code.

# Heading Refresher

## Subheading

Fix the subheading above so it shows up correctly.

### Sub Sub Heading

# Spacing

If you would like text on two lines make sure there are two spaces after.  
Like this.


# Aspect Ratio, Alignment, and Captions

Look at the knitted version. Then add dpi= 250, fig.width=7, fig.height=7 to the code chunk header and knit again. Adding fig.cap with add a caption and fig.align can be changed to 'center', 'left' or 'right'

```{r, dpi= 250, fig.width=7, fig.height=7, fig.cap = "*Fig.1 Diamonds Graph*", fig.align='center'}
diamonds %>% 
      ggplot(aes(x = color, fill = color)) +
      geom_bar() +
      facet_grid(clarity ~ cut) +
      labs(title = "Title", 
          subtitle = "Subtitle") + 
      theme_new + theme(legend.position = "right", 
                        legend.direction = "vertical", 
                        legend.key.size = unit(.3, "cm"))
```

# Run all above and below.

Find these buttons and play with them.

# Visual Markdown Editor

Find this buttons and play with them.

# Knitting to doc versus html
 
Q: How do you knit to a word doc?  
A: You just click the word doc knit button. But you can't have any interactive features in your file. 

Q: Can you edit a knitted word doc?  
A: Yes you can, but it will be overwritten by knitting again. I would recommend only doing this as a very last step. 

# How to write good comments 

Add comments to the following code chunk
```{r}
# Creating a new copy of the diamonds dataset
diamonds1 <- diamonds

# Creating a table 
diamonds1 %>% 
  filter(table >= 55) %>% # that shows just observations with a table of greater than 55
  group_by(color) %>% 
  summarize(med_price = median(price)) %>% # And the median price for each color
  arrange(desc(med_price)) %>% 
  head(5) # Only outputting the top 5

# Overwriting the table variable to be integers
diamonds1 <- diamonds1 %>% mutate(table = round(table))
```

# Tables 
Take a look at the gross output of the code chunk above. We can use the pander package to make this a lot more appealing in our knitted file.  
```{r}
diamonds1 %>% 
  filter(table >= 55) %>% # that shows just observations with a table of greater than 55
  group_by(color) %>% 
  summarize(`median price` = median(price)) %>% # And the median price for each color
  arrange(desc(`median price`)) %>% 
  head(5) %>% # Only outputting the top 5 
  pander()
```

You will be outputting 1 summarization table in your knitted report. 

# Inline code

Put the mean in the following text.

The mean of the price variable is `r mean(diamonds$price, na.rm = TRUE)`.

# Debugging 
Uncomment the code below and try to debug it.  
```{r}
diamonds1 %>% 
  mutate(price_and_color = paste0("This diamond of color ", str_to_lower(color), " is $", format(round(price), big.mark = ",")),
         table = round(table)) %>% head() %>% pander()
```

The following tips may help: 

1. Little red x 
2. Rainbow parentheses 
3. Stack overflow 
4. eval = FALSE


# Projects
Create a dummy project. 
```{r}
getwd()
```

Q: Why would you use a relative or absolute file path?  
A: If you are ever sharing code with someone it is a good idea to have it in a project, and then use relative file paths within the project so that they can run the exact same code on their computer without having to edit. 

# RStudio Hygiene 

1. Clean the global environment regularly
2. Restart whenever you have a new session. (OR when you have a code chunk that was working before but isn't now!)
3. Create a new session if you are working on multiple different projects at the same time. 


# Practice 
How would I run this code in my markdown file but make sure it doesn't output anything? 
```{r, include = FALSE}
summary(diamonds)
```

How would I keep it from running at all? 
```{r, eval = FALSE}
summary(diamonds)
```

How would I run it but hide the code? 
```{r, echo = FALSE}
summary(diamonds)
```

How would I hide any warnings that come in the code? 
```{r, warning = FALSE}
summary(diamonds)
```