---
title: "HW.Clean.5"
author: "Logan Hardy"
date: "2/28/2022"
output: html_document
---

25 points (20 for this assignment, 5 for being on time)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
library(tidyverse)
library(lubridate)
library(ggthemes)
```


# Import (2 points)
Import the executive comp dataset and the Stolen_Vehicles_2012_Salt_Lake_City dataset with readr. Import all date time columns in a date-time format. Strings should be factors. 
```{r}
exec_comp_main <- read_csv("../Data/exec_comp_main.csv", col_types = cols(Formation.yr = col_datetime(format = "%Y"))) %>% mutate_if(is.character, as.factor)

Stolen_Vehicles_2012_Salt_Lake_City <- read_csv("../Data/Stolen_Vehicles_2012_Salt_Lake_City.csv", col_types = cols(`REPORT DATE` = col_datetime(format = "%m/%d/%Y %I:%M:%S %p"), `OCC DATE` = col_datetime(format = "%m/%d/%Y %I:%M:%S %p"))) %>% mutate_if(is.character, as.factor)

```

# Outliers 

## Finding outliers with summary (3 points)
*In the console* run a summary of the exec_comp_main dataset. 

Q: Name at least 3 columns that you think have outliers and describe why you think so.  
A: Revenue - mean is significantly larger than median and 3rd quartile (indicating its right skewed and probably has outliers)
Expenses - mean is significantly larger than median and 3rd quartile (indicating its right skewed and probably has outliers)
Liabilities - mean is significantly larger than median and 3rd quartile (indicating its right skewed and probably has outliers)
Assets - mean is significantly larger than median and 3rd quartile (indicating its right skewed and probably has outliers)
Reportable.comp - Max is significantly further away from mean/median than min
Health.and.benefits - Mean is significantly larger than median and 3rd quartile (indicating its right skewed and probably has outliers) 



## Finding outliers with visualizations (2 points)
Make a histogram and boxplot of the Liabilities column of the exec_comp_main dataset.
```{r}
theme_new <-  theme_fivethirtyeight(base_size=12#,base_family="Open Sans"
                                    ) %+replace% theme(panel.grid.major.y = element_line(colour = "grey80", size = 0.25),  panel.grid.major.x = element_blank(), panel.background = element_rect(fill = "white"), plot.background = element_rect(fill = "white"), legend.background = element_rect(fill = "white")) 

exec_comp_main %>% ggplot(aes(x = Liabilities)) + geom_histogram(fill = "#afd7db", bins = 10) +
        labs(title = "Histogram of Liabilities", 
            subtitle = "Notice how the max greatly pulls the graph to the right and there should be \n10 bins, but the other counts are too small to see", 
            y = "Count", 
            x = "Liabilities") +
        theme_new #+ theme(axis.title = element_text(size = 10, color = "grey40"))

exec_comp_main %>% ggplot(aes(x = Liabilities)) + 
        geom_boxplot(fill = "#afd7db") +
        labs(title = "Boxplot of Liabilities", 
            subtitle = "Note how the box can't be seen because of outliers", 
            x = "Liabilities") +
        theme_new #+ theme(axis.title = element_text(size = 10, color = "grey40"))

```

Q: Based on these graphs, why or why not do you think there are outliers in this variable?  
A: Based on these graphs, I think that there are outliers because neither graph fully shows.  

## Using the Boxplot function to find outliers (3 points)
Use the boxplot function to find the index of outliers in the Expenses column. Output the head() of a subset of just rows with outliers in that variable. Then output the head() of a subset that has a new column that says whether there are outliers in that expenses column. 
```{r}
boxplot_outliers <- boxplot.stats(exec_comp_main$Expenses)$out
outlier_indexes <- which(exec_comp_main$Expenses %in% c(boxplot_outliers))

# dataset of just outliers. 
exec_comp_main[outlier_indexes, ] %>% head()

# subset that has a new column that says whether there are outliers in that expenses column. 
exec_comp_main %>% 
  mutate(ExpensesOutlier = ifelse(Expenses %in% c(boxplot_outliers), TRUE, FALSE)) %>% head()

```


Q: How many rows have an outlier in the expenses column according to the boxplot function?  
A: 7,021

## Using percentile to replace outliers with NAs (3 points)
Make a copy of the exec_comp_main dataset. Create a cut off of 95% and 5% for outliers. Replace all outliers in the Assets colum with NA. 
```{r}
exec_comp_main_NAs <- exec_comp_main
  
lower_bound <- quantile(exec_comp_main_NAs$Assets, 0.05, na.rm = TRUE)
upper_bound <- quantile(exec_comp_main_NAs$Assets, 0.95, na.rm = TRUE)

# This provides the indexes of outliers
outlier_indexes <- which(exec_comp_main_NAs$Assets < lower_bound | exec_comp_main_NAs$Assets > upper_bound)

# And makes a dataset with just outliers
exec_comp_main_NAs[outlier_indexes, "Assets"] <- NA

# Count NAs
#sum(is.na(exec_comp_main$Assets))
#sum(is.na(exec_comp_main_NAs$Assets))

```

Q: How many NAs were in that column before you replaced outliers?  
A: 0

Q: How many NAs are in the dataset copy with replaced NAs?  
A: 4969

# Date-Time
## Creating new date-time columns (4 points)
All the functions and formatting we did with date times can be used within a mutate function to create new date-time columns. 

Create a new copy of the stolen vehicle dataset and overwrite it to include 4 new columns: 

1. a column called REPORTING_QUARTER that has the year and quarter when the incident was reported
2. a column called REPORTING_LAPSE that has the difference between the time the incident happened (OCC DATE) and when it was reported in days (hint: if you nest each column in as.Date they will output the difference in days not seconds.)
3. a column called INCIDENT_WEEKDAY that has the weekday (written out, like "Sunday") when the incident happened
4. a column called INCIDENT_WEEK that has a numerical value for which week in the year the incident happened. (Sunday should be the first day of that week.)

Then output the head() of the dataset. 
```{r}
library(zoo)
Stolen_Vehicles_2012_Salt_Lake_City_COPY <- Stolen_Vehicles_2012_Salt_Lake_City

Stolen_Vehicles_2012_Salt_Lake_City_COPY <- Stolen_Vehicles_2012_Salt_Lake_City_COPY %>% mutate(REPORTING_QUARTER <- as.yearqtr(`REPORT DATE`, "%m/%d/%Y %I:%M:%S %p"), REPORTING_LAPSE <- (as.Date(`REPORT DATE`) - as.Date(`OCC DATE`)), INCIDENT_WEEKDAY <- format(`OCC DATE`, "%A"), INCIDENT_WEEK <- format(`OCC DATE`, "%U")) 

Stolen_Vehicles_2012_Salt_Lake_City_COPY %>% head()

```

## Other date-time practice (3 points)
```{r}
# Turn the following string into a POSIXct format
as.POSIXct("June 14, 22 (Tuesday) 10:34 pm", format="%B %d, %y (%A) %I:%M %p")


# Turn the following string into a POSIXct format (assume it is in correct Mountain Daylight Time), then convert it to the time zone of Central Europe (CET).
date1 <- as.POSIXct("Feb 20 2022 13:55:22", format="%b %d %Y %H:%M:%S")
with_tz(date1, tzone = "CET")

# Turn the following string into a date format then extract the month fully spelled out.  
date2 <- as.Date("2022-11-06", format="%Y-%m-%d")
format(date2, "%B")

```


--End--

# Optional feedback: 

How hard was this assignment? Was it: too hard, too easy, or just right? 

ANSWER: Just Right 

Approximately how long did this assignment take you? 

ANSWER: 3 hours 

