---
title: "Transform.2 In Class Starter (dplyr 2)"
author: "Logan Hardy"
date: "01/20/2022"
output: html_document
---


# Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Import the refugee2011_15 dataset

```{r}
refugees2011_15 <- read_csv("../Data/refugees2011-15.csv") %>% mutate_if(is.character, as.factor)
```

# Arrange

The arrange function is the primary way we sort our dataset. We can arrange alphabetically or numerically. We can also sort by multiple variables at the same time and by descending order.

Q: How is the data currently sorted? 
A: 

```{r}
# Sort the data by Country of origin name in alphabetical order
refugees2011_15 %>% arrange(Origin)

# Sort the data by Country of origin in reverse alphabetical order
refugees2011_15 %>% arrange(desc(Origin))


# Sort the data by the Total population in ascending order. 
refugees2011_15 %>% arrange(TotalPop)

# Sort the data by refugees in descending order.
refugees2011_15 %>% arrange(desc(Refugees))


# Sort the data by Country of Origin, then by descending asylyum seekers. 
refugees2011_15 %>% arrange(Origin, desc(AsylumSeekers))

```

Q: In the following code, which sorting order takes precedence? refugees2011_15 %>% arrange(Country, desc(Origin), Year) 
A: Country 

Q: In the same code above, How is a tie of country and origin settled? (Aka what happens if more than one row has the same Country and Origin?) 
A: Year 

## Using arrange and head together

Arrange and head are commonly used together because sometimes we just want to see certain information.

```{r}
# Grab only the 10 rows with the highest refugee count
refugees2011_15 %>% arrange(desc(Refugees)) %>% head(n=10)

```

Q: Filtering also enables us to grab a certain number of rows, just like head() and arrange() do together. What is the difference between using these two approaches? 
A: 

```{r}
# Use fiter to return the same subset as the code above (It doesn't need to be ordered the same.)
refugees2011_15 %>% filter(Refugees >= 951142)

```

# Summarize

Summarize is used to collapse numeric columns into some sort of summary statistic. That means it always works in conjunction with some sort of mathematical function, like mean(), median(), sum(), etc. It will only return the columns it directly pertains to.

```{r}
# Find the sum of the entire Asylum column. 
refugees2011_15 %>% summarise(sum(AsylumSeekers, na.rm = TRUE))

```

Q: How does summarize default to naming the column it creates? 
A: It names it after the function used to create it.

This autocreated name is kind of ugly. We will always want to rename it.

```{r}
# Repeat the same code above but this time name it total_asy
refugees2011_15 %>% summarise(total_asy = sum(AsylumSeekers, na.rm = TRUE))

```

# Group_by

Summarizing is useful to a point, but more often than wanting to know the summary statistic of an entire column (which we could do just as easily by using the sum function by itself), we will want to know a summary statistic for each category in a categorical variable.

If we group our dataset by a categorical variable first, then use the summarize function, it will return one row for every category with the summary statistic.

```{r}
#Find the mean number of refugees coming from every origin country. Remember to rename the new column.
refugees2011_15 %>% 
  group_by(Origin) %>% 
  summarize(mean_refugee = mean(Refugees, na.rm = TRUE), sum_ref = sum(Refugees, na.rm = TRUE))
```

Q: What happens if you do group_by() without a summarize function? 
A: 

Q: What happens if you try to summarize a category that doesn't have any data in it? 
A: 

Just like the arrange() function, you can layer groupings to group by more than one categorical variable at once. You can also summarize more than one numeric variable at once. 

```{r}
# Find the maximum refugee count for every origin and asylum combination.
refugees2011_15 %>% group_by(Origin, Country) %>% summarise(max_ref = max(Refugees, na.rm = TRUE))

# Use that same code, but this time include a column for the maximum asylum seeker count. 
refugees2011_15 %>% group_by(Origin, Country) %>% summarise(max_ref = max(Refugees, na.rm = TRUE), max_asylum = max(AsylumSeekers, na.rm = TRUE)) 
```

Q: Does it make sense to use group_by() and summarize() separately? 
A: 

Q: Does the group_by() come before or after the summarize? 
A: 

## Counting groups
We said that when we use group_by and summarize together we are finding a statistic about a numeric variable based on a categorical variable. There is one major exception to this rule. Sometimes we just want to know how many rows are in each category. In this case we don't need to involve any numeric variable. 

```{r}
# Does each year have the same number of data points? 
refugees2011_15 %>% group_by(Year) %>% summarise(count = n())
```

Q: We said we group_by categorical variables. But isn't year a numerical column? 
A: 

```{r}
# Which Origin and Country combination have the highest number of rows?  
refugees2011_15 %>% group_by(Origin, Country) %>% summarise(count = n()) %>% arrange(desc(count)) %>% head(1)
```

## A quirk of group_by
What happens when you group by a variable but then want to select other variables? 
```{r}
# group the data by the year variable then select only the origin and country columns. 
refugees2011_15 %>% group_by(Origin, Country) %>% select(Year, Refugees)
```

Solving this problem: the ungroup() function 
```{r}
refugees2011_15 %>% group_by(Origin, Country) %>% ungroup() %>% select(Year, Refugees)

```

# Putting it all together 
```{r}
# I want to know which 5 countries have been sought asylum from the US between 2011-2015


# Make a new column that subtracts asylum seekers from refugees. Total this number for every year. Make a new column that says "in 2011, there were x more refugees than asylum seekers." Call this column output. Select only this column as the eventual output. 


# Count the number of asylum countries with more than 100,000 collective refugees in 2013 and 2015


# Find the standard deviation of refugees and asylum seekers that come from each country between 2011-2015. (I want to see a separate column for refugees and asylum seekers) Remove all missing data. Return all rows with a standard deviation of each measure above or equal to 1000. Sort your results first by the standard deviation of refugees in descending order, then the standard deviation of asylum seekers in descending order. Only include the top 5 results. 

```



