---
title: "Explore.1 In class starter"
author: "Logan Hardy"
date: "01/06/2022"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

# Update your name and the date at the top of this document.

# Introduction to RStudio

Copy and paste the following code lines into the console one at a time.

random_number <- 3 
random_number + random_number 
plot(1:5)

Look at the four quadrants in RStudio.

Q: How do you change the size of the panes?
A: Drag edges to desired size

Q: What is the purpose of each pane?
A: R script file, command console, environment/history/tutorial, outputs

# R Markdown vs. Script

Copy and paste the code lines from above into the code chunk below and run them.

This is an R Markdown File (RMD or .rmd).
Open a script (.r) and see the difference.

```{r}
# {r, echo=FALSE, eval=FALSE, message=FALSE, include=FALSE}
random_number <- 3
random_number + random_number 
plot(1:5)
```

Q: What are 3 differences between Markdown Files and scripts?
A:

# Objects, Vectors, and Lists

You can create both numerical objects and text objects.
Create one of each below.

```{r}
words <- "garlic bread"
number <- 24
```

Saved text is usually referred to as a "string" in R.
Can you perform mathematical operations on strings?
Try this in the console.

Create two vectors, one numerical and one text.
Print them out.

```{r}
numbers <- c(2:24)
#numbers <- c(3,1,6)
numbers

(strings <- c("hat", "lemon", "3"))


```

Create a blank vector

```{r}
blank <- c()
nas <- c(NA)

```

Create a vector and print it out at the same time.

```{r}
(numbers2 <- c(1:10))
```

Print out the second value in your string vector.

```{r}
strings[2]
```

Combine two vectors together.

```{r}
strings3 <- c("hi", "family")
(strings4 <- c(strings, strings3))

boolean1 <- c(TRUE, FALSE, TRUE)
TRUE + TRUE
```

Make a list.

```{r}
list1 <- list(TRUE, 2, "Sandwich")
```

Make a list within a list.
Click on it in the global environment.

```{r}
list2 <- list(list1, "nice", "class")
```

Access a value in the nested list with its index

```{r}
list1[3]
list2[[1]][3]
```

Make a named list

```{r}
list3 <- list("num1" = 23, "num2" = 1, "num3" = 2)
list3
```

Access part of a list by its name

```{r}
list3$num2

list3[["num2"]]
list3[[2]]
```

Q: What is the difference between a vector and a list?
A: vectors are one dimensional and one datatype.
Lists can be multidimensional and can store multiple datatypes

# Datasets

Mtcars is a dataset that is always loaded into R.
Run the code below to see it.

```{r}
mtcars
```

Q: What is the structure of a dataset?
A: A named list of vectors

Print out the first column of the mtcar dataset by its index.

```{r}
mtcars[1]
```

Print out the first column of the mtcars dataset by its name.

```{r}
mtcars$mpg
```

Save the first value of that vector as an object

```{r}
first_value <- mtcars$mpg[1]
first_value
```

Make a vector called "name" that stores the names of your family members.

```{r}
name <- c("Logan", "Katelyn", "Thomas", "Jori", "Olivia")
name
```

Make a vector called "age" that stores the ages of those family members in the same order.

```{r}
(age <- c(24, 25, 2, 1, 1))
```

Make a dataset called "family" that stores these two columns.
Print it out.

```{r}
family <- data.frame(name, age)
family
```

Find the mean of the ages of your family

```{r}
meanAge <- mean(family$age)
meanAge
```

Knit this document and look at the output.
